|

### 2019-04-29 rap/article.Rmd
typos
add text

---
title:  |
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author: 
  - "Danilo Freire[^df]"
  - "John Meadowcroft[^jm]"
  - "David Skarbek[^ds]"
  - "Eugenia Guerrero[^eg]"
date: \today
thanks: "We thank Fábio Barros, Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments. Replication materials and R source code are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
abstract: "This article presents a georeferenced dataset on human rights violations in the Pinochet dictatorship in Chile. We coded the personal details of 2,398 victims named in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations. The dataset comprises 57 variables from 1973 to 1990 and is available as a stand-alone spreadsheet or as the `pinochet` package for `R`. As examples, we describe the major temporal and spatial patterns of the human rights abuses. We also discuss our coding procedures, show descriptive statistics, and conclude with suggestions for potential applications of the dataset."
keywords: "Chile, georeferenced event data, human rights, Pinochet, truth commission"
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    template: template.latex
---

@@

On 11 September 1973, General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende. The coup marked the beginning of a seventeen-year military dictatorship which combined rapid economic liberalisation with large-scale political repression [@meadowcroft2014hayek; @valdes1995pinochet]. In 1991, after Chile's transition to democracy, President Patricio Aylwin convened a commission to investigate human rights violations in the Pinochet regime. The Report of the National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], recorded over 2,000 cases of murders and disappearances from 1973 to 1990. The Commission confronted the abuses of the past and portrayed a careful reconstruction of state violence during the military government. Most importantly, the Report contributed to the erosion of authoritarianism in Chile, as it fostered a series of criminal investigations that culminated with the indictment of General Pinochet in 2000 [@brahm2007uncovering 26].

@@

Violence during transitions has characterised regime change in Latin America. Transitions to autocratic rule in Brazil (1964) and Argentina (1976) prompted frequent human rights violations that only recently have been fully documented [@agosin1987generals; @elpais2016argentina]. In some cases, autocrats declared unrestrained violence to be an explicit government policy. In 1992, Peru's former President Alberto Fujimori openly stated that his _autogolpe_ granted him the rights to "[apply] drastic punishments" towards any suspect of terrorism [@dabene1997narcodemocracias, 97; @samtleben2013constitucion, 56].

@@

The establishment of the Truth and Reconciliation Commission was the first attempt to make the dictatorship accountable for its crimes. The Commission focused on the most serious cases of human rights violations, those which resulted in deaths of disappearances[^omissions]. One factor that limited the scope of the investigations was that the Commission did not have the legal authority to demand collaboration from the military forces [@popkin1995truth, 85]. The Chilean Commission negotiated an amnesty to perpetrator in exchange for freedom of inquiry, and it collected testimony from victims in Chile and abroad [@quinn2001dealing, 399]. The result was a broad yet accurate description of the events. Although the military contested the conclusions of the Report, they did not question its factual information [@bakiner2009denial, 54; @brahm2005beyond].

@@

\vspace{.5cm}
\begin{center}
Table 1 about here.

Example entry with the victim's characteristics\\ and information about their murder or disappearance.
\end{center}
\vspace{.5cm}

[comment]: <> (Example Entry)

@@

@@

\vspace{.5cm}
\begin{center}
Table 2 about here.

Number of killings, disappearances, unresolved cases, and suicides 
\end{center}
\vspace{.5cm}

[comment]: <> (Table 2 around here)

@@

Two time trends: one with the whole period,\\ another discarding the first year to bette visualise variation in later years
\end{center}
\vspace{.5cm}

[comment]: <> (Figure 02 - Time Trends Here)
\newpage

@@

[comment]: <> (Figure 03 - Spatial variation)

@@

[comment]: <> (Figure 04 - Students and Communist Leaders)

>>>>

---
title:  |
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author: 
  - "Danilo Freire[^df]"
  - "John Meadowcroft[^jm]"
  - "David Skarbek[^ds]"
  - "Eugenia Guerrero[^eg]"
date: \today
thanks: "We thank Umberto Mignozzetti and Robert Myles McDonnell for their helpful comments. Replication materials and R source code are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
abstract: "This article presents a georeferenced dataset on human rights violations in the Pinochet dictatorship in Chile. We coded the personal details of 2,398 victims named in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations. The dataset comprises 57 variables from 1973 to 1990 and is available as a stand-alone spreadsheet or as the `pinochet` package for `R`. As examples, we describe the major temporal and spatial patterns of the human rights abuses. We also discuss our coding procedures, show descriptive statistics, and conclude with suggestions for potential applications of the dataset."
keywords: "Chile, georeferenced event data, human rights, Pinochet, truth commission"
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    template: template.latex
---

@@

On 11 September 1973, General Augusto Pinochet led a coup against Chile's socialist President Salvador Allende. The coup marked the beginning of a seventeen-year military dictatorship which combined rapid economic liberalisation with large-scale political repression [@valdes1995pinochet]. In 1991, after Chile's transition to democracy, President Patricio Aylwin convened a commission to investigate human rights violations in the Pinochet regime. The Report of the National Commission on Truth and Reconciliation [-@report1991], also known as the Rettig Report[^rettig], recorded over 2,000 cases of murders and disappearances from 1973 to 1990. The Commission confronted the abuses of the past and portrayed a careful reconstruction of state violence during the military government. Most importantly, the Report contributed to the erosion of authoritarianism in Chile, as it fostered a series of criminal investigations that culminated with the indictment of General Pinochet in 2000 [@brahm2007uncovering 26].

@@

Violence during transitions has characterised regime change in Latin America. Transitions to autocratic rule in Brazil (1964) and Argentina (1976) prompted frequent human rights violations that only recently have been fully documented [@agenciabrasil2014ditadura; @elpais2016argentina]. In some cases, autocrats declared unrestrained violence to be an explicit government policy. In 1992, Peru's former President Alberto Fujimori openly stated that his _autogolpe_ granted him the rights to "[apply] drastic punishments" towards any suspect of terrorism [@dabene1997narcodemocracias, 97; @samtleben2013constitucion, 56].

@@

The establishment of the Truth and Reconciliation Commission was the first attempt to make the dictatorship accountable for its crimes. The Commission focused on the most serious cases of human rights violations, which are those that resulted in deaths of disappearances[^omissions]. One factor that limited the scope of the investigations was that the Commission did not have the legal authority to demand collaboration from the military forces [@popkin1995truth, 85]. The Chilean Commission negotiated an amnesty to perpetrator in exchange for freedom of inquiry, and it collected testimony from victims in Chile and abroad [@quinn2001dealing, 399]. The result was a broad yet accurate description of the events. Although the military contested the conclusions of the Report, they did not question its factual information [@bakiner2009denial, 54; @brahm2005beyond].

@@

| Variable                | Sample          |
| :---------------------- | :-------------- |
| region                  | Taparacá        |
| id                      | 276             |
| page                    | 265             |
| start date              | 18/01/1974      |
| end date                | 18/01/1974      |
| last name               | Marques Agusto  |
| first name              | Nelson José     |
| minor                   | 0               |
| age                     | 31              |
| male                    | 1               |
| occupation              | Office worker   |
| victim affiliation      | Communist party |
| violence                | 1 (killed)      |
| method                  | 1 (gun)         |
| interrogation           | NA (unknown)    |
| torture                 | NA (unknown)    |
| mistreatment            | 1 (yes)         |
| perpetrator             | Military        |
| place                   | 5 (in public)   |
| start location          | Iquique         |
| latitude 1              | -20.23070       |
| longitude 1             | -70.13567       |
| place 2                 | 4 (in custody)  |
| location 2              | Pisagua prison  |
| latitude 2              | -19.59667       |
| longitude 2             | -70.21222       |

Table: Sample event from the dataset (selected variables).

@@

| Violence         | _N_    |
| :--------------- | -----: |
| Killed           |   1331 |
| Disappeared      |    853 |
| Unresolved cases |     93 |
| Suicides         |      7 |
| Unknown          |        |

Table: Number of killings, disappearances, unresolved cases, and suicides

@@

@@

Two time trends: one with the whole period,\\ another discarding the first year to better visualise variation in later years
\end{center}
\vspace{.5cm}

[comment]: <> (Figure 01 - Time Trends Here)
\newpage

@@

[comment]: <> (Figure 02 - Spatial variation)

@@

[comment]: <> (Figure 03 - Students and Communist Leaders)
